---
title: "Bactch Import"
author: "Baboyma Kagniniwa"
date: "2022-06-22"
output: html_notebook
---

## Load Packages and Activate Environment

```{r}
.rs.restartR()

library(reticulate)

env <- "test-hyper"
#env <- "test-venv"

use_virtualenv(virtualenv = env, required = TRUE)

py_config()

```

## Generate Test Processed Files

Find and Read Latest MSD Global PSNUxIM file

```{r}
library(tidyverse)
library(vroom)
library(gophr)
library(glamr)

source("../Scripts/utilities.R")

dir_mer <- si_path() %>% paste0("../", .)
dir_out <- "../Dataout"
dir_csv <- dir_out %>% paste0(., "/MSD/PSNUxIM")

file_psnu_im <- return_latest(folderpath = dir_mer, 
                              pattern = "MER_.*_PSNU_IM_FY20-.*")

df_psnu <- file_psnu_im %>% read_msd()

```

Identify Fiscal Years & Operating Units

```{r}
fys <- df_psnu %>% distinct(fiscal_year) %>% pull() %>% sort()
ous <- df_psnu %>% distinct(operatingunit) %>% pull %>% sort
```

Extract Orgunits Reference Data set

```{r}
df_orgs <- df_psnu %>% 
  select(operatingunit:dreams) %>% 
  distinct()

file_orgs <- paste0("MSD_PSNU_IM_FY20-22_Orgunits_", 
                    str_extract(file_psnu_im, "\\d{8}"), ".csv")

df_orgs %>% 
  write_csv(x = ., file = file.path(dir_csv, file_orgs), na = "")
```

Extract Mechanisms Reference Data set

```{r}
file_mechs <- paste0("MSD_PSNU_IM_FY20-22_Mechanisms_", 
                     str_extract(file_psnu_im, "\\d{8}"), ".csv")

df_mechs <- df_psnu %>% 
  select(funding_agency, mech_code, mech_name, 
         award_number, starts_with("prime")) %>% 
  distinct()

df_mechs %>% 
  write_csv(x = ., file = file.path(dir_csv, file_mechs), na = "")
```

Extract Data Elements

```{r}
file_de <- paste0("MSD_PSNU_IM_FY20-22_DataElements_", 
                     str_extract(file_psnu_im, "\\d{8}"), ".csv")

df_de <- df_psnu %>% 
  select(fiscal_year, indicator:modality, source_name) %>% 
  distinct() 

df_de %>% 
  write_csv(x = ., file = file.path(dir_csv, file_de), na = "")
```

Extract FY/OU Program Raw Data sets

```{r}
walk(ous, function(.x){
  
  print(paste0("OU: ", .x))
  
  walk(fys, function(.y){
    
    print(paste0("FY: ", .y))
  
    file_out <- paste0("MSD_PSNU_IM_FY", 
                       str_sub(.y, 3, 4), "_", 
                       str_replace(.x, "'", ""), "_", 
                       str_extract(file_psnu_im, "\\d{8}"), ".csv")
    
    print(file_out)
      
    df_ou_fy <- df_psnu %>% 
      filter(operatingunit == .x, fiscal_year == .y) 
    
    if (nrow(df_ou_fy) > 0) {
      df_ou_fy %>% 
        select(fiscal_year, psnuuid, mech_code, 
               indicator:categoryoptioncomboname,
               targets:last_col()) %>% 
        #reshape_msd() %>% 
        #mutate(fiscal_year = paste0("20", str_sub(period, 2, 4))) %>% 
        write_csv(x = ., file = file.path(dir_csv, file_out), na = "")
    }
  })
})

```

## Read Reference and Processed files into a Database (Hyperfile)

Retrieve list of files to be ingested

```{r}
# list reference files
ref_orgs_file <- list.files(path = dir_csv, 
                            full.names = TRUE,
                            pattern = "MSD_PSNU_IM_FY20-22_Org.*_\\d{8}.csv$")

ref_mechs_file <- list.files(path = dir_csv, 
                             full.names = TRUE,
                             pattern = "MSD_PSNU_IM_FY20-22_Mech.*_\\d{8}.csv$")

ref_de_file <- list.files(path = dir_csv, 
                          full.names = TRUE,
                          pattern = "MSD_PSNU_IM_FY20-22_DataE.*_\\d{8}.csv$")

# list files with FY22 Data
proc_fy22_files <- list.files(path = dir_csv, 
                              full.names = TRUE,
                              pattern = "MSD_PSNU_IM_FY22_.*_\\d{8}.csv$")

# list all files
proc_files <- list.files(path = dir_csv, 
                         full.names = TRUE,
                         pattern = "MSD_PSNU_IM_FY\\d{2}_.*_\\d{8}.csv$")

# output file
output_file = file_psnu_im %>% 
  basename() %>%
  str_replace(".zip$", ".hyper") %>% 
  paste0("../Dataout/Hypers/", .)

```

Read csv files with python

Note that `R` variables / objects can be accessed from `python` with `r.`. `py$` is used for the reverse.

```{python}
import pandas as pd
df = pd.read_csv(r.ref_orgs_file)

df.head()
```

Use python functions through modules

```{python}
#import os
#os.path.abspath(os.path.join(".."))
```

```{python}
import os
from time import sleep
import pandas as pd
import pantab as pt
from tableauhyperapi import TableDefinition, TableName, SqlType

modules_path = os.path.abspath(os.path.join(".."))

if modules_path not in sys.path:
  sys.path.append(modules_path + "\\Scripts")
  
import utilities as util


output_file = r.output_file

util.prinf("Processing ...")

util.prinf([output_file])

# Orgs

util.prinf(r.ref_orgs_file)

df_orgs = util.pd.read_csv(r.ref_orgs_file)
#df_orgs = util.csv_to_df(r.ref_orgs_file)

pt.frame_to_hyper(df_orgs, database=output_file, table="Orgunits")

# Mechs

util.prinf(r.ref_mechs_file)

df_mechs = util.pd.read_csv(r.ref_mechs_file)
#df_mechs = util.csv_to_df(r.ref_mechs_file)

pt.frame_to_hyper(df_mechs, database=output_file, table="Mechanisms", table_mode="a")

# DataElements

util.prinf(r.ref_de_file)

df_de = util.pd.read_csv(r.ref_de_file)
#df_mechs = util.csv_to_df(r.ref_de_file)

pt.frame_to_hyper(df_de, database=output_file, table="DataElements", table_mode="a")

# Data Table Definition
# tbl_data = TableDefinition(
#   TableName("msd", "PSNUxIM"),
#   [
#     TableDefinition.Column("fiscal_year", SqlType.text()),
#     TableDefinition.Column("psnuuid", SqlType.text()),
#     TableDefinition.Column("mech_code", SqlType.text()),
#     TableDefinition.Column("indicator", SqlType.text()),
#     TableDefinition.Column("numeratordenom", SqlType.text()),
#     TableDefinition.Column("indicatortype", SqlType.text()),
#     TableDefinition.Column("disaggregate", SqlType.text()),
#     TableDefinition.Column("standardizeddisaggregate", SqlType.text()),
#     TableDefinition.Column("categoryoptioncomboname", SqlType.text()),
#     TableDefinition.Column("targets", SqlType.text()),
#     TableDefinition.Column("qtr1", SqlType.text()),
#     TableDefinition.Column("qtr2", SqlType.big_int()),
#     TableDefinition.Column("qtr3", SqlType.big_int()),
#     TableDefinition.Column("qtr4", SqlType.big_int()),
#     TableDefinition.Column("cumulative", SqlType.big_int()),
#     TableDefinition.Column("source_name", SqlType.text())
#   ]
# )

tbl_data_types = {
    "fiscal_year": "str",
    "psnuuid": "str",
    "mech_code": "str",
    "indicator": "str",
    "numeratordenom": "str",
    "indicatortype": "str",
    "disaggregate": "str",
    "standardizeddisaggregate": "str",
    "categoryoptioncomboname": "str",
    "targets": "str",
    "qtr1": "str",
    "qtr2": "float",
    "qtr3": "float",
    "qtr4": "float",
    "cumulative": "float",
    "source_name": "str"
}

tbl_data = "PSNUxIM"


print(f"Number of files #{len(r.proc_files)}")

# load files 1 at the time
for idx, ifile in enumerate(r.proc_files):

  #print(f"Input file: {ifile}")
  print(f"Input file #{idx}: {ifile}")

  # Read csv files with pre-defined column data types
  df_msd = pd.read_csv(ifile, dtype = tbl_data_types)
  
  # Mutate data structure
  
  # Wide format
  # df_msd = df_msd.astype(tbl_data_types)
  
  # Long format
  # df_msd.loc[:, df_msd.columns != "value"].astype('str')
  # df_msd.loc[:, df_msd.columns == "value"].astype('float')
  
  # load file content to hyperfile through append mode
  try:
    pt.frame_to_hyper(df = df_msd, database=output_file, table=tbl_data, table_mode="a")
  except Exception as e:
    print(e)
    continue
  
print("Done!")
  
```


