---
title: "Using R and Python to Create Tableau Extract"
author: "Baboyma Kagniniwa"
date: "2022-06-15"
output: html_notebook
---

This is a sample code used to test [Tableau API](https://www.tableau.com/developer/tools/hyper-api) to generate [Tableau Extract](https://help.tableau.com/current/pro/desktop/en-us/extracting_data.htm) as `hyperfile`. This process leverages R and Python through `reticulate` R package located [here](https://rstudio.github.io/reticulate/). 

Credits to [RStudio Team](https://rstudio.github.io/reticulate/articles/rstudio_ide.html), [Alexander](https://datastorydesign.com/python/tableau/2020/06/17/read-and-write-hyper-files-in-python.html), and stockoverflow for all the tips.

## Configuration

The configurations should be run once - except for when installing new packages / libraries.

```{r setup, include=F, eval=F}
# Install and load reticulate package
# This will also prompt you to install r-miniconda

install.packages("reticulate")

# Environment name
env <- "test-venv"

# Create a virtual environment => env will located in ~/.virtualenvs/
# Make sure to version of python matches the one in miniconda
reticulate::virtualenv_create(envname = env, version = "3.8.10")

# Activate venv
reticulate::use_virtualenv(virtualenv = env)

# Check configs
reticulate::py_available()
reticulate::py_version()
reticulate::py_config()

# Install Python Libraries
reticulate::py_install(package = "pandas", envname = env)
reticulate::py_install(package = "tableau-api-lib", envname = env)
reticulate::py_install(package = "tableauhyperapi", envname = env)
reticulate::py_install(package = "tableauhyperio", envname = env)

```

## Load Packages and Activate Environment

```{r}
library(reticulate)

env <- "test-venv"

use_virtualenv(virtualenv = env, required = TRUE)

```

## Check configs

```{r}
# Make sure the response is TRUE
py_available()
```

```{r}
# Confirm miniconda python version
py_version()
```

```{r}
# Confirm configurations
py_config()
```

## Load sample data

```{r warning=F, message=F}
library(tidyverse)
library(nycflights13)

readr::write_csv(flights, file = "../Data/flights.csv", na = "")
```

## Read and filter data with python / pandas

```{python}
import pandas as pd

flights = pd.read_csv("../Data/flights.csv")

flights = flights[flights["dest"] == "ORD"]
flights = flights[['carrier', 'dep_delay', 'arr_delay']]
flights = flights.dropna()

flights.head()

```

## Explore filtered data with R

```{r}
str(py$flights)
head(py$flights)
```

## Visualize output data

```{r}
ggplot(py$flights, aes(carrier, arr_delay, color = carrier)) +
  geom_point(show.legend = F) + 
  geom_jitter(show.legend = F)
```

## Export output as Tableau Extract

```{python}
# Load TBL Hyper IO Library
import tableauhyperio as hio

# Output file name 
output_file = "../Dataout/flights_output.hyper"

# Writing a regular hyper file
hio.to_hyper(flights, output_file)

```

## Validate Tableau Extract

```{python}

# Reading data from regular hyper file
df = hio.read_hyper(output_file)

df.head()

```

## Source Python code from R

```{r}
source_python("../Scripts/hello.py")
```
